---
title: "AD-nonAD-AsymAD Grouping and Comparison of Diagnosis Strategies for AD"
output: html_document
date: '2022-12-21'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Call related packages from library
```{r}
library(readxl)
library(tidyverse)
library(limma)

setwd("C:\\path of file")
getwd()
```

Normalized and covariate adjusted RNA sequencing data from the ROSMAP study by Neff et al. (Neff et al., 2021) for DLPFC brain region was downloaded from the Synapse database (dataset ID of syn25944586). 
```{r}
table_exp<-read_excel("data.xlsx")
rownames(table_exp)<-table_exp$EnsemblID

data<- table_exp[,3:616]
rownames(data)<- table_exp$EnsemblID

# This file includes metadata of ROSMAP downloaded from Synapse. https://www.synapse.org/#!Synapse:syn3157322
info <- read_excel("file_name.xlsx")
```

Individuals were grouped based on clinical score (cogdx=4|5 for AD group, cogdx=1|2|3 for non-AD group)
```{r}
control_cogdx <- info %>% filter (cogdx==1 | cogdx==2| cogdx==3) 
control_cogdx_exp<- data %>% select (control_cogdx$Sample_id)
rownames(control_cogdx_exp)<- table_exp$EnsemblID

AD_cogdx<- info %>% filter (cogdx==4 | cogdx==5)  # | cogdx==6 was not included because of other dementia
AD_cogdx_exp<- data %>% select (AD_cogdx$Sample_id)
rownames(AD_cogdx_exp)<- table_exp$EnsemblID

data_cog<- cbind(control_cogdx_exp,AD_cogdx_exp)
rownames(data_cog)<- table_exp$EnsemblID
```

Individuals were grouped based on neuropathological score (CERAD=1|2 for AD group, cogdx=3|4 for non-AD group)
```{r}
score6<- c("Proj10292311", "Proj11455530", "Proj15495024", "Proj20179164" ,"Proj93815598", "Proj50300408", "Proj10405008", "Proj50108598", "Proj68015667","Proj20201189","Proj37527863")

control_cerad <- info %>% filter (CERAD==3 | CERAD==4) %>% filter(!Sample_id %in% score6) # These samples were removed because of other dementia.
control_cerad_exp<- data %>% select (control_cerad$Sample_id)
rownames(control_cerad_exp)<- table_exp$EnsemblID

AD_cerad<- info %>% filter (CERAD==1 | CERAD==2) %>% filter(!Sample_id %in% score6)
AD_cerad_exp<- data %>% select (AD_cerad$Sample_id)
rownames(AD_cerad_exp)<- table_exp$EnsemblID

data_cer<- cbind(control_cerad_exp,AD_cerad_exp)
rownames(data_cer)<- table_exp$EnsemblID
```


Consensus non-AD and consensus AD between cogdx and CERAD.
Determining AsymAD group.
```{r}
control_common <- info %>% filter (CERAD==3 | CERAD==4) %>% filter (cogdx==1 | cogdx==2| cogdx==3)%>%filter(!Sample_id %in% score6)

control_common_exp<- table_exp %>% select (control_common$Sample_id)
rownames(control_common_exp)<- rownames(table_exp)

AD_common<- info %>% filter (CERAD==1 | CERAD==2) %>% filter (cogdx==4 | cogdx==5) 
AD_common_exp<- table_exp %>% select (AD_common$Sample_id)
rownames(AD_common_exp)<- rownames(table_exp)


# AsymAD group includes 91 individuals.
AD91<- info %>% filter (cogdx==1)%>% filter (CERAD==1 | CERAD==2)
AD91_exp<- table_exp %>% select (AD91$Sample_id)
rownames(AD91_exp)<- rownames(table_exp)
```

DEGs were listed with limma package.
non-AD vs AD based on cogdx
non-AD vs AD based on CERAD
non-AD vs AD based on consensus categorization
two additional categorization strategies
For five differetn categorization strategies DEGs were determined and recorded.
```{r}

#This is an example for determination of DEGs for cogdx based categorization

designMat <- c(rep('control', 359), rep('AD', 243)) # 
design<-model.matrix(~0 + factor(designMat))
colnames(design)<-c('control','AD')
contrast <- makeContrasts(AD - control, levels = design)
fit <- lmFit(data_cog, design)
 fit2 <- contrasts.fit(fit, contrast)
  fit2 <- eBayes(fit2)
  degDF1 <- as.data.frame(topTable(fit2, n = 'Inf'))

write.csv(degDF1,"DEGs_cogdx.csv")
```

