---
title: "metobolom"
output: html_document
date: '2022-11-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Metabolome Data Analysis
```{r}
library(readxl)
library(tidyverse)
library(SuperExactTest)
setwd("C:\\path")
getwd()
```


Metabolome data is available for some of the individuals in the ROSMAP cohort, collected using the untargeted metabolomics platform from Metabolon Inc (Metabolon HD4, syn25985690). Metabolites with >30 % missing values were excluded.  For each remaining metabolite, the missing values were imputed as half of the minimum measured value of that metabolite. 

```{r}
data<- read_excel("data.xlsx")
datat<- as.data.frame(t(data))
data_clear<- datat[-c(1:5),]
colnames(data_clear)<- datat[2,]
data_clear<- as.data.frame(sapply(data_clear, function(x) as.numeric(x)))
met<- rownames(datat)[6:691]
rownames(data_clear)<- met
my_sample<- colnames(data_clear)
data_clear$minimum <-apply(data_clear, 1, function(x) min(x, na.rm = TRUE)/2)
data_pro<- data_clear
for(i in 1:nrow(data_pro)){
  na_rows <- grepl('NA',data_pro[i,])
  data_pro[i,na_rows] <- data_pro$minimum[i]
}
```

The data was adjusted for the gender, age of death, and post-mortem interval of the samples using the lm function in R.
```{r}
metadatarosmap <- read_excel("metadatarosmap.xlsx")
filtered_meta<- metadatarosmap[metadatarosmap$projid %in% my_sample, ]
data_pro<- data_pro[,1:240]
data_clear<- data_pro

data_d<- as.data.frame(t(data_clear)) 

condition.num = filtered_meta$condition
sex.num = filtered_meta$msex
age= as.data.frame(filtered_meta$age_death)
age[age=="90+"] = "91"
age=as.numeric(age$`filtered_meta$age_death`)
pmi= as.numeric(filtered_meta$pmi)

fit3 = lm(as.matrix(data_d)~ sex.num + age + condition.num + pmi) 

coef=fit3[["coefficients"]]

colnames(coef)
data_d2= t(data_d)

fitteddata=matrix(0, 686,240)
cins=sex.num
cins[cins=="female"]=0
cins[cins=="male"]=1
cins= as.numeric(cins)
state=condition.num
state[state=="AD"]=0
state[state=="control"]=1
state=as.numeric(state)
names=NULL
for (j in 1:240){
   
   for (i in 1:686) {
      genename= colnames(coef)[i]
      con= state[j]
      gender= cins[j]
      deathage= age[j]
      pmitime=pmi[j]
      count= data_d2[i,j]-gender*coef[2,i]-deathage*coef[3,i]-pmitime*coef[5,i]
      names[i]= genename
      fitteddata[i,j]= count
   }
}
colnames(fitteddata)= colnames(data_d2)
rownames(fitteddata)=names
write.csv(fitteddata,"covariat_Adjust_metabolon_mindivided2.csv")
save(fitteddata,file="results.RData")
```

Read data
```{r}
info <- read_excel("file_name.xlsx")
metabolon<- read.csv("covariat_Adjust_metabolon_mindivided2.csv",header = TRUE, sep=";")
```


Grouping samples similar to transcriptome data analysis
```{r}
#For clinical categorization
control_cogdx <- info %>% filter (cogdx==1 | cogdx==2| cogdx==3) %>% filter(.$Sample_id %in% colnames(metabolon))
control_cogdx_exp<- metabolon %>% select (control_cogdx$Sample_id)
rownames(control_cogdx_exp)<- metabolon$X

AD_cogdx <- info %>% filter (cogdx==4 | cogdx==5) %>% filter(.$Sample_id %in% colnames(metabolon)) 
AD_cogdx_exp<- metabolon %>% select (AD_cogdx$Sample_id)
rownames(AD_cogdx_exp)<- metabolon$X

data_cogdx<- cbind(control_cogdx_exp,AD_cogdx_exp)
rownames(data_cogdx)<-metabolon$chemical

#For pathological categorization
control_cerad <- info %>% filter (CERAD==3 | CERAD==4) %>% filter(.$Sample_id %in% colnames(metabolon))
control_cerad_exp<- metabolon %>% select (control_cerad$Sample_id)
rownames(control_cerad_exp)<- metabolon$X

AD_cerad <- info %>% filter (CERAD==1 | CERAD==2) %>% filter(.$Sample_id %in% colnames(metabolon))
AD_cerad_exp<- metabolon %>% select (AD_cerad$Sample_id)
rownames(AD_cerad_exp)<- metabolon$X

data_cerad<- cbind(control_cerad_exp,AD_cerad_exp)
rownames(data_cerad)<-metabolon$chemical

#For consensus categorization
control_common <- info %>% filter (CERAD==3 | CERAD==4) %>% filter (cogdx==1 | cogdx==2| cogdx==3)  %>% filter(.$Sample_id %in% colnames(metabolon))
control_common_exp<- metabolon %>% select (control_common$Sample_id)

AD_common<- info %>% filter (CERAD==1 | CERAD==2) %>% filter (cogdx==4 | cogdx==5)%>%  filter(.$Sample_id %in% colnames(metabolon)) 
AD_common_exp<- metabolon %>% select (AD_common$Sample_id)

consensus<- cbind(control_common_exp,AD_common_exp)
rownames(consensus) <- metabolon$chemical

#For AsymAD group
AD41<- info %>% filter (cogdx==1)%>% filter (CERAD==1 | CERAD==2)%>% filter(.$Sample_id %in% colnames(metabolon))
AD41_exp<- metabolon %>% select (AD41$Sample_id)
rownames(AD41_exp)<- metabolon$chemical

# For AsymAD-adjusted clinical categorization
control_cogdx_adj<- select(control_cogdx_exp,-AD41$Sample_id)
AD_cogdx_adj<- cbind(AD41_exp,AD_cogdx_exp)
data_adjusted1<-cbind(control_cogdx_adj,AD_cogdx_adj)
rownames(data_adjusted1)<- metabolon$chemical

# For AsymAD-adjusted pathological categorization
control_cerad_adj<- cbind(control_cerad_exp,AD41_exp)
AD_cerad_adj<- select(AD_cerad_exp,-AD41$Sample_id)
data_adjusted2<-cbind(control_cerad_adj,AD_cerad_adj)
rownames(data_adjusted2)<- metabolon$chemical
```

Classical t-test was used.
```{r}
# It is an example for cogdx based categorization.
pval <- vector()
for (i in 1:nrow(data_cogdx)) {  
  pval[i] = t.test(as.numeric(data_cogdx[i,1:134]), as.numeric(data_cogdx[i,135:234]))$p.value
}
pval<- as.data.frame(pval)
pval_res<-cbind(metabolon$chemical,pval)

write.csv(pval_res,"metabolon_pval_cogdx_noscore6_ttest.csv")
```

super exact test
https://cran.r-project.org/web/packages/SuperExactTest/vignettes/set_html.html
https://cran.r-project.org/web/packages/SuperExactTest/SuperExactTest.pdf
```{r}
data<- read_excel("super_exact_data.xlsx")
p=cpsets(8,c(83,48,60),686,lower.tail=FALSE)
print(p) # 7.874683e-10

x=list(pat=data$metabolit[1:83], cli=data$metabolit[84:131], cons=data$metabolit[132:191])

fit=MSET(x, 686, FALSE)
fit$p.value  #2.062149e-08

x_pc=list(pat=data$metabolit[1:83], cons=data$metabolit[132:191]) #compare pathologic vs consensus p= 3.2x10-20

x_cc=list(cli=data$metabolit[84:131], cons=data$metabolit[132:191])#compare clinic vs consensus, p=3.9x10-14

x_patcli=list(pat=data$metabolit[1:83], cli=data$metabolit[84:131])#compare clinic vs pathologci, p=0.12

fit=MSET(x_patcon, 686, FALSE)

Result=supertest(x, n=686)
plot(Result,sort.by='size')
```
